The imprint of global magma oceans on exoplanet demographics

### 2023-03-16 src/tex/ms.tex
add a robustness test for the statistical estimator

\bibliographystyle{aasjournal}
\bibliography{bib,PhD,coauthors}
\end{document}

>>>>

\appendix
\section{Robustness tests}
\subsection{Alternative statistics for the average radius or bulk density}\label{app:binnedstats}
The hypothesis tests introduced in Sect.~\ref{sec:met_surveys-hypotests} rely on a statistical estimator for the variation of planetary radii or bulk densities as a function of net instellation, and we chose a moving average for this estimator in our nominal setup.
Here, we explore how robust our results are against this choice by demonstrating the recovery of the runaway greenhouse signal in the case of the optimistic survey (Sect.~\ref{sec:res_testability}) with an alternative estimator: instead of computing moving averages, we used a binned statistic.

We first binned the data of the simulated survey in instellation space, choosing the number of bins via the rule of \citet{Freedman1981} and using logarithmic binning.
In each bin, we computed the arithmetic mean of the planet radius and its standard deviation.
Then, we assigned each planet the mean radius according to the instellation bin it occupies and used this as the measure for testing the runaway greenhouse hypothesis.

\begin{figure}[ht!]
%    \script{figurescript.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/optimistic_R-S_binned.pdf}
        \caption{
            Detection of the runaway greenhouse threshold with a binned statistic.
            Using simulated data from the optimistic survey case ($N = \var{N_optimistic}$, compare Fig.~\ref{fig:optimistic_R-S}), we tested the runaway greenhouse hypothesis. Instead of using a running mean, we computed a binned statistic (arithmetic mean and standard deviation in blue) to assign planets the average radius or bulk density in their neighborhood in instellation space. Random draws from the posterior of the runaway greenhouse hypothesis (Eqn.~\ref{eq:rgh_hypo}) are shown in red.
            As in the nominal case, the pattern is detected with high significance.
        }
        \label{fig:optimistic_R-S_binned}
    \end{centering}
\end{figure}
Figure~\ref{fig:optimistic_R-S_binned} shows the simulated data together with binned, average planet radii and draws from the posterior of the hypothesis test.
A clear detection resulted, although with somewhat lower significance ($\Delta \ln \mathcal{Z} \approx 30$) compared to the nominal setup.
The accuracy of the recovered instellation threshold is comparable.
This test demonstrates that our results are not sensitive on the choice of statistical estimator to test the hypotheses against.

It is conceivable that with very large sample sizes and a very sharp runaway greenhouse transition a binned solution would perform better.
For a search with real data, both approaches, and possibly other alternatives, should be considered.

\bibliographystyle{aasjournal}
\bibliography{bib,PhD,coauthors}

\end{document}
